---
title: "Statistics_solutions"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Statistics_solutions}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


```{r}
library(tidyverse)
library(broom)
library(openintro)
```

## Exercise

1. Check whether there is a difference in the performance at baseline (`ohs0`) between the males and females. Study the distributions and decide which test to use. 

2. We used an ANOVA to study the distribution of satisfaction.cat and age. Check the distribution of age over the 3 satisfaction.cat groups. Does it pass normality assumptions? If you are unsure, use Kruskal-Wallis test. Hint you can get information using `?kruskal.test`.

3. Use ANOVA or Kruskal-Wallis test to check if there is an association between baseline ohs0 and satisfaction.cat groups.

4. Create a table of retired against satisfaction groups. Does the chi-square test show an association between the two? 

```{r}
hip<-read.csv("../inst/extdata/hip_data.csv")

hip %>% filter(!is.na(sex)) %>% ggplot(aes(ohs0, fill=sex)) + facet_grid(~sex) + geom_histogram()


w<-wilcox.test(ohs0~sex,data=hip %>% filter(!is.na(sex)))
w

hip<-hip %>% mutate(satisfaction.cat=factor(ifelse(satisfaction<80,"Low",ifelse(between(satisfaction,80,90),"Medium","High")), levels=c("Low","Medium","High")))

hip %>% filter(!is.na(satisfaction.cat)) %>% ggplot(aes(age, fill=satisfaction.cat)) + facet_grid(~satisfaction.cat) + geom_histogram()


kr <- hip %>% filter(!is.na(satisfaction.cat)) %>% 
kruskal.test(age~satisfaction.cat, data=. )

table(hip$retired, hip$satisfaction.cat)

chisq.test(hip$retired, hip$satisfaction.cat)

```



1\. Fit a new model that uses  `pf_expression_control` to predict `hf_score`, or the total human freedom score. What does the slope tell us in the context of the relationship between human freedom and the amount of political pressure on media content?


2\. How does this relationship compare to the relationship between `pf_score` and `pf_expression_control`?
    Use the $R^2$ values from the two model summaries to compare.
    Does your `pf_expression_control` seem to predict `hf_score` better?
    Why or why not?

3\. Check the model diagnostics using appropriate visualisations and evaluate if the model conditions have been met.

4\. Pick another pair of variables of interest and visualise the relationship between them.
    Do you find the relationship surprising or is it what you expected.
    Discuss why you were interested in these variables and why you were/were not surprised by the relationship you observed.

```{r}
hfi_2016 <- hfi %>% 
  filter(year == 2016)

m1 <- lm(hf_score ~ pf_expression_control, data = hfi_2016)
m2 <- lm(pf_score ~ pf_expression_control, data = hfi_2016)

summary(m1)
#


ggplot(data = hfi_2016, aes(x = pf_expression_control, y = hf_score)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE)


ggplot(data = hfi_2016, aes(x = pf_expression_control, y = pf_score)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE)

plot(m1)

```




## EXERCISES



1\. Add region to the model from Q1 using `lm(hf_score ~ pf_expression_control + region, data = hfi_2016)`. What do you notice about the slope between human freedom and expression control? How has it changed from the previous model. Do you think region is a confounder, and think about reasons why this might be so? 

2\. Compare the $R^2$ for the 2 models from Q1 and the previous questions. Is there an increase by adding region? Think about the definition of $R^2$. What does this mean in this context?

3.\ Fit a new model that uses  `ef_money` or monetary measure to predict `hf_score`.  What does the slope tell us in the context of the relationship between human freedom and the economy in the country? 

4\. Again add region to the model from Q3. Compare the slope and $R^2$ with the model from Q3.

6\. Finally fit a model with `ef_money` and `pf_expression_control` as exposures and `hf_score` as outcome.  Compare the slope and $R^2$ from the models from the previous exercises. Could `ef_money` be a confounder? 

7\. Use a linear regression model (and scatter plot) with  `ef_money` as exposure and `pf_expression_control` as outcome, to study whether `ef_money` has an association with `pf_expression_control` as well. This might validate our finding that `ef_money` is a confounder between  `pf_expression_control` as exposure and 
`hf_score` as outcome from Q6.


```{r}
tidy(m1)
m2 <- lm(hf_score ~ pf_expression_control + region, data = hfi_2016)
summary(m1)
summary(m2)

m3 <- lm(ef_money ~ pf_expression_control, data = hfi_2016)
m4 <- lm(ef_money ~ pf_expression_control + region, data = hfi_2016)

m1 <- lm(hf_score ~ pf_expression_control + ef_money, data = hfi_2016)

m4 <- lm(hf_score ~  ef_money, data = hfi_2016)


ggplot(data = hfi_2016, aes(x = ef_money, y = pf_expression_control)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE)
m5 <- lm(pf_expression_control ~ ef_money, data = hfi_2016)


```



